Measuring individual differences in the understanding of gaze cues across the lifespan

### 2021-12-21 paper/manuscript.Rmd
add first sentences for methods

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

@@

## Material

## Procedure

@@

>>>>

abstract: |
  There must be an abstract of no more than 250 words.
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

@@

## Material
Material is presented as an interactive web-app that is accessible for computers and tablets and runs on any web browser.
The code is open-source (https://github.com/ccp-eva/gafo-demo) 
and a live demo version can be found under: https://ccp-odc.eva.mpg.de/gafo-demo/.

The web-app generates two files: 
(1) a text file (JSON) containing click responses, 
and (2) a video file (webm) of the participant's webcam recording.
Data gets automatically collected and safely stored on local servers located in Leipzig, Germany.

The web-app was programmed in Vanilla JavaScript, HTML and CSS. 
For stimulus presentation, a scalable vector graphic (SVG) composition was parsed.
This way, the composition scales according to the user's viewport without loss of quality,
while keepiong the aspect ratio and relative object positions.
Furthermore, SVGs allow us to precisely control and calculate the size and position of all 
composite parts of the scene (e.g., pupil of the agent).

The GreenSock Animation Platform (GSAP; TODO insert citation) library was used 
to animate the movement of single SVG elements.

Study instructions were pre-recorded and, therefore, constant across participants. 
No interaction with the experimenter was necessary and children needed minimal assistance from their caregivers.

## Procedure (from pre-reg)
An animal character (i.e., agent; sheep, monkey, or pig) is placed centrally in a window. 
A balloon (i.e., target; blue, green, yellow, or red) is located in front of them. 
The target then falls to the ground. 
The agent’s gaze follows the movement of the target, 
that is, the pupils of the agent move in a way that their center aligns with the center of the target. 
Participants are then asked to locate the target’s position on the screen.

Visual access to the target’s true location is manipulated by a hedge. 
Participants either have full, partial, or no visual access to the true target location. 
When partial or no information about the target location is accessible, 
participants are expected to use the agent’s gaze as a cue.

There are two different versions of partial and no visual access trials. 
For families that use a tablet with touchscreen, 
the target lands behind a hedge (i.e., hedge version). 
Children are then asked to click directly on the hedge to indicate where the target is.
For families that use a computer without touchscreen, the balloon falls into a box. 
Children are then asked to point to the box containing the target. 
Caregivers then act as the "digital finger" of their children and click on the indicated box.

The dependent variable depends on the study version: 
For the hedge version, our dependent variable is continuous. 
Here, the dependent variable is imprecision, 
which is defined as the absolute difference between the true x coordinate of the target 
and the x coordinate of the participant’s click on the screen. 
For the box version, we use our categorical outcome (i.e., which box was clicked) 
to calculate the proportion of correct responses.

There are three different trial types, 
depending on the visual access to the target flight and end location. 
In touch training trials, participants have full visual access to the target flight 
and the target's end location. 
In familiarization trials, participants have visual access to the target flight 
but not to the target's end location. 
In the hedge version, the target's end location is covered by a hedge. 
In the box version, the target flies into a box and is therefore not visible anymore. 
In test trials, participants have no visual access to the target flight 
nor the end location. 
In both versions, the target flight is covered by a hedge. 
In the hedge version, the hedge then shrinks to cover the target's end location. 
In the box version, the hedge shrinks completely. 
The boxes then hide the target's end location.

Depending on the participant's device, 
participants have to indicate their estimated target location directly on the hedge 
(i.e., hedge version) or in one of five boxes (i.e., box version).

Participants receive 19 trials with one touch training, two familiarization trials, 
and 16 test trials. 
The first trial of each type comprises a voice-over description of the presented events. 
Touch and familiarization trials, as well as voice-over trials, will not be included in the analysis. 
Therefore, we will conduct our analysis with 15 test trials.

Target locations are generated as follows. 
For the hedge version, the full width of the screen is divided into ten bins. 
Exact coordinates within each bin are randomly generated. 
For the box version, the target randomly lands in one out of five boxes. 
Each bin/box occurs equally often. 
The same bin/box can occur only twice in a row.

@@

# Declarations
## Open practices statement
The web application (https://ccp-odc.eva.mpg.de/gafo-demo/) described here is open source (https://github.com/ccp-eva/gafo-demo).
The datasets generated during and/or analysed during the current study are available
in the [gazecues-methods] repository, (https://github.com/jprein/gazecues-methods).
All experiments were preregistered (https://osf.io/zjhsc/).

## Funding 
This study was funded by the Max Planck Society for the Advancement of Science, a noncommercial, publicly financed scientific organization (no grant number). 
We thank all the children and parents who participated in the study.

## Conflicts of interest 
The authors declare that they have no conflict of interest.

## Ethics approval

## Consent to participate
Informed consent was obtained from all individual participants included in the study or their legal guardians.

## Consent for publication

## Open access

## Authors' contributions 
optional: please review the submission guidelines from the
journal whether statements are mandatory

### 2021-12-21 paper/manuscript.tex
add first sentences for methods

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{Measuring individual differences in the understanding of gaze cues}
\author{Julia Prein\textsuperscript{1}, Manuel Bohn\textsuperscript{1}, Luke Maurits\textsuperscript{1}, Steven Kalinke\textsuperscript{1}, \& Daniel M. Haun\textsuperscript{1}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Measuring individual differences in the understanding of gaze cues},
  pdfauthor={Julia Prein1, Manuel Bohn1, Luke Maurits1, Steven Kalinke1, \& Daniel M. Haun1},
  pdflang={en-EN},
  pdfkeywords={social cognition, individual differences, gaze cues, psychometrics},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\abstract{
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

@@

@@

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{Measuring individual differences in the understanding of gaze cues across the lifespan}
\author{Julia Prein\textsuperscript{1}, Manuel Bohn\textsuperscript{1}, Luke Maurits\textsuperscript{1}, Steven Kalinke\textsuperscript{1}, \& Daniel M. Haun\textsuperscript{1}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Measuring individual differences in the understanding of gaze cues across the lifespan},
  pdfauthor={Julia Prein1, Manuel Bohn1, Luke Maurits1, Steven Kalinke1, \& Daniel M. Haun1},
  pdflang={en-EN},
  pdfkeywords={social cognition, individual differences, gaze cues, psychometrics},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\abstract{
There must be an abstract of no more than 250 words.
One or two sentences providing a \textbf{basic introduction} to the field, comprehensible to a scientist in any discipline.

@@

Idea for an opener :)

Developmental psychology is facing a dilemma: many research questions are questions about individual differences, yet, there is a lack of tasks to reliably measure these individual differences. For example \ldots{} .

@@

Material is presented as an interactive web-app that is accessible for computers and tablets and runs on any web browser.
The code is open-source (\url{https://github.com/ccp-eva/gafo-demo})
and a live demo version can be found under: \url{https://ccp-odc.eva.mpg.de/gafo-demo/}.

The web-app generates two files:
(1) a text file (JSON) containing click responses,
and (2) a video file (webm) of the participant's webcam recording.
Data gets automatically collected and safely stored on local servers located in Leipzig, Germany.

The web-app was programmed in Vanilla JavaScript, HTML and CSS.
For stimulus presentation, a scalable vector graphic (SVG) composition was parsed.
This way, the composition scales according to the user's viewport without loss of quality,
while keepiong the aspect ratio and relative object positions.
Furthermore, SVGs allow us to precisely control and calculate the size and position of all
composite parts of the scene (e.g., pupil of the agent).

The GreenSock Animation Platform (GSAP; TODO insert citation) library was used
to animate the movement of single SVG elements.

Study instructions were pre-recorded and, therefore, constant across participants.
No interaction with the experimenter was necessary and children needed minimal assistance from their caregivers.

\hypertarget{procedure-from-pre-reg}{%
\subsection{Procedure (from pre-reg)}\label{procedure-from-pre-reg}}

An animal character (i.e., agent; sheep, monkey, or pig) is placed centrally in a window.
A balloon (i.e., target; blue, green, yellow, or red) is located in front of them.
The target then falls to the ground.
The agent's gaze follows the movement of the target,
that is, the pupils of the agent move in a way that their center aligns with the center of the target.
Participants are then asked to locate the target's position on the screen.

Visual access to the target's true location is manipulated by a hedge.
Participants either have full, partial, or no visual access to the true target location.
When partial or no information about the target location is accessible,
participants are expected to use the agent's gaze as a cue.

There are two different versions of partial and no visual access trials.
For families that use a tablet with touchscreen,
the target lands behind a hedge (i.e., hedge version).
Children are then asked to click directly on the hedge to indicate where the target is.
For families that use a computer without touchscreen, the balloon falls into a box.
Children are then asked to point to the box containing the target.
Caregivers then act as the ``digital finger'' of their children and click on the indicated box.

The dependent variable depends on the study version:
For the hedge version, our dependent variable is continuous.
Here, the dependent variable is imprecision,
which is defined as the absolute difference between the true x coordinate of the target
and the x coordinate of the participant's click on the screen.
For the box version, we use our categorical outcome (i.e., which box was clicked)
to calculate the proportion of correct responses.

There are three different trial types,
depending on the visual access to the target flight and end location.
In touch training trials, participants have full visual access to the target flight
and the target's end location.
In familiarization trials, participants have visual access to the target flight
but not to the target's end location.
In the hedge version, the target's end location is covered by a hedge.
In the box version, the target flies into a box and is therefore not visible anymore.
In test trials, participants have no visual access to the target flight
nor the end location.
In both versions, the target flight is covered by a hedge.
In the hedge version, the hedge then shrinks to cover the target's end location.
In the box version, the hedge shrinks completely.
The boxes then hide the target's end location.

Depending on the participant's device,
participants have to indicate their estimated target location directly on the hedge
(i.e., hedge version) or in one of five boxes (i.e., box version).

Participants receive 19 trials with one touch training, two familiarization trials,
and 16 test trials.
The first trial of each type comprises a voice-over description of the presented events.
Touch and familiarization trials, as well as voice-over trials, will not be included in the analysis.
Therefore, we will conduct our analysis with 15 test trials.

Target locations are generated as follows.
For the hedge version, the full width of the screen is divided into ten bins.
Exact coordinates within each bin are randomly generated.
For the box version, the target randomly lands in one out of five boxes.
Each bin/box occurs equally often.
The same bin/box can occur only twice in a row.

@@

\hypertarget{declarations}{%
\section{Declarations}\label{declarations}}

\hypertarget{open-practices-statement}{%
\subsection{Open practices statement}\label{open-practices-statement}}

The web application (\url{https://ccp-odc.eva.mpg.de/gafo-demo/}) described here is open source (\url{https://github.com/ccp-eva/gafo-demo}).
The datasets generated during and/or analysed during the current study are available
in the {[}gazecues-methods{]} repository, (\url{https://github.com/jprein/gazecues-methods}).
All experiments were preregistered (\url{https://osf.io/zjhsc/}).

\hypertarget{funding}{%
\subsection{Funding}\label{funding}}

This study was funded by the Max Planck Society for the Advancement of Science, a noncommercial, publicly financed scientific organization (no grant number).
We thank all the children and parents who participated in the study.

\hypertarget{conflicts-of-interest}{%
\subsection{Conflicts of interest}\label{conflicts-of-interest}}

The authors declare that they have no conflict of interest.

\hypertarget{ethics-approval}{%
\subsection{Ethics approval}\label{ethics-approval}}

\hypertarget{consent-to-participate}{%
\subsection{Consent to participate}\label{consent-to-participate}}

Informed consent was obtained from all individual participants included in the study or their legal guardians.

\hypertarget{consent-for-publication}{%
\subsection{Consent for publication}\label{consent-for-publication}}

\hypertarget{open-access}{%
\subsection{Open access}\label{open-access}}

\hypertarget{authors-contributions}{%
\subsection{Authors' contributions}\label{authors-contributions}}

optional: please review the submission guidelines from the
journal whether statements are mandatory
